---
title: "London Housing Price Over the Years 2009-2019"
author: "A.Mayuri(2348133)"
date: "2024-02-15"
output: word_document
---
## Question:

Choose a real time series data and illustrate the following questions.
1. Perform the various steps to understand about the components of the data and comment on
it.
2. Discuss the mathematical model for the given data.
3. Comment about the behavior of Autocorrelation function plot.
4. Does the logarithmic transformation helps in achieving stationarity of the data set? Justify
your answer.
Briefly prepare a report based on the above questions.

## Objective:


1) To Understand the components of the data
2)Develop a time series model 
3)comment on the ACF plot
4)Lograthimic transformation needed to achieve stationarity?


## Dataset:
The dataset has been sourced from https://www.kaggle.com/datasets/justinas/housing-in-london
where the price of houses has been recorded monthly over the 11 years period. from 2009-2019.

## Data Description:

1)location(london): this is uniform throught out the dataset which confirms that this is a time series data-set and not a panel/cross-sectional data.

2) Date: The dataset is monthly data thus the monthly dates are recorded. with corresponding monthly price averages.

3) Price average: Price average is computed by summing over all the prices by the number of houses sold in that particular minth

##Exploratory Data Analyses: 
Since the 
1)dataset taken has the same observation ie (Place-london),code(city=E92000001) 
2)the date being a non disruptive/ no gaps and hence continuous in nature.
3) Average price of the houses (Variable of interest)
Performing a complete time series analyses is the best way to understand the data (EDA).

### Import Dataset

```{r }
library(readr)
LHPC <- read_csv("C:/Users/mayur/Desktop/Mstat/Semesters/Tri-sem3/Time series/Dataset/LHPC.csv")
View(LHPC)
attach(LHPC)
```

### Import data set -(Time series - only price component)
```{r }
library(readr)
LHP1 <- read_csv("C:/Users/mayur/Desktop/Mstat/Semesters/Tri-sem3/Time series/Dataset/LHP1.csv")
View(LHP1)
attach(LHP1)
```

### Convert data into Time-Series And Plot
```{r }
data2=ts(LHP1,start=2009,frequency=12) # converting it into a time series data
data2
LHP1
length(LHP1)
ts.plot(data2)
```

Interpretation: Here we observe that the dataset only has a trend (upward) and irregularity component. we cannot use a multiplicative model due to the absence of seasonality component.

Thus, the model is of additive form with trend and irregularity component.

#### MATHEMATICAL MODEL:
                        z(t)=m(t)+e(t)
                        where, z(t)-time series variable dependent on time,
                        m(t)-trend component
                        e(t)-irregularity component.
                        
### Auto-Correlation Function Plot:

```{r }
acf(data2)
```

Here the model has not achieved stationarity because the the acf lines are well above the band line. The conversion into stationary process is demonstrated using differencing method below.


## Lograthimic Transformation And Stationarity

1) We cannot use lograthimic transformation since it helps us to convert a multiplicative model into an additive model.

2) Since our model is additive we need not use lograthimic transformation.

3) To achiev estationarity we need to use differencing metod as demonstrated below.


## stationary classes of time series model. (Justification)

we need to transform the data to stationary form. a polynomial degree say p form model will give a stationary model post differencing p times.

### differencing method and Plotting (Demonstration)
```{r }

diffdata=diff(data2)
diffdata
ts.plot(diffdata)
library(tseries)

```
which is a stationary process, 
since it has a 
1)constant variance 
2)Constant mean.


## Confirm the stationarity (Augmented Dicky -Fuller Test)
h0: the data trend is not stationary
h1: the data trend is stationary

```{r }
adf.test(diffdata)




```
since p value is less than the significance value alpha(0.05). Thus we reject null hypothesis and conclude that the data is stationary in nature.



```{r }

```



```{r }

```




```{r }

```



```{r }

```



```{r }

```


```{r }

```


```{r }

```


```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```


```{r }

```
