---
title: "London Housing Price Over the Years 2009-2019"
author: "A.Mayuri(2348133)"
date: "2024-02-15"
output: word_document
---
## Question:

Choose a non seasonal time series data,
1)illustrate the method of differencing.
2)ordinary least squares 
3)moving average smoothing to extract the stationary version.


## Objective:
To Extract Stationary component(Part) from Non-Stationary Data

1)Perform Basic operations to understand the dataset
2)Differencing Method
3)Ordinary Least Square Method
4)Moving Average Smoothing Method.


## Dataset:
The dataset has been sourced from https://www.kaggle.com/datasets/justinas/housing-in-london
where the price of houses has been recorded monthly over the 11 years period. from 2009-2019.

## Data Description:

1)location(london): this is uniform throught out the dataset which confirms that this is a time series data-set and not a panel/cross-sectional data.

2) Date: The dataset is monthly data thus the monthly dates are recorded. with corresponding monthly price averages.

3) Price average: Price average is computed by summing over all the prices by the number of houses sold in that particular minth

##Exploratory Data Analyses: 
Since the 
1)dataset taken has the same observation ie (Place-london),code(city=E92000001) 
2)the date being a non disruptive/ no gaps and hence continuous in nature.
3) Average price of the houses (Variable of interest)
Performing a complete time series analyses is the best way to understand the data (EDA).

### Import Dataset

```{r }
library(readr)
LHPC <- read_csv("C:/Users/mayur/Desktop/Mstat/Semesters/Tri-sem3/Time series/Dataset/LHPC.csv")
View(LHPC)
attach(LHPC)
```

### Import data set -(Time series - only price component)
```{r }
library(readr)
LHP1 <- read_csv("C:/Users/mayur/Desktop/Mstat/Semesters/Tri-sem3/Time series/Dataset/LHP1.csv")
View(LHP1)
attach(LHP1)
```

### Convert data into Time-Series And Plot
```{r }
data2=ts(LHP1,start=2009,frequency=12) # converting it into a time series data
data2
LHP1
length(LHP1)
ts.plot(data2)
```

Interpretation: Here we observe that the dataset only has a trend (upward) and irregularity component. we cannot use a multiplicative model due to the absence of seasonality component.

Thus, the model is of additive form with trend and irregularity component. 

Note: The data has no seasonality component involved.

#### MATHEMATICAL MODEL:
                        z(t)=m(t)+e(t)
                        where, z(t)-time series variable dependent on time,
                        m(t)-trend component
                        e(t)-irregularity component.
                        
### Auto-Correlation Function Plot:

```{r }
acf(data2)

```

Here the model has not achieved stationarity because the the acf lines are well above the band line. The conversion into stationary process is demonstrated using differencing method below.

## Method Of Differencing. 

#### Concept: we need to transform the data to stationary form. a polynomial degree say p form model will give a stationary model post differencing p times.

### differencing method and Plotting (Demonstration)
```{r }

diffdata=diff(data2)
diffdata
ts.plot(diffdata)
library(tseries)

```
which is a stationary process, 
since it has a 
1)constant variance 
2)Constant mean.


## Confirm the stationarity (Augmented Dicky -Fuller Test)
h0: the data trend is not stationary
h1: the data trend is stationary

```{r }
adf.test(diffdata)




```
since p value is less than the significance value alpha(0.05). Thus we reject null hypothesis and conclude that the data is stationary in nature.

## Moving Average Smoothing:

#### Concept: 

Moving averages are a series of averages calculated using sequential segments of data points over a series of values. They have a length, which defines the number of data points to include in each average

Types

Centered Moving Averages: It include both previous and future observations to calculate the average at a given point in time.

One Sided Moving Average: 
One-sided moving averages include the current and previous observations for each average. For example, the formula for a moving average (MA) of X at time t with a length of 5 is the following:

MA(5)=X(t-4)+X(t-3)+X(t-2)+X(t-1)+X(t)/5

```{r }
library(forecast)
ma1=ma(data2, order=3)
ma2=ma(data2,order=5)
ma3=ma(data2,order=7)
ma3=ma(data2,order=9)
par(mfrow=c(2,2))# combine multiple plots in one
ts.plot(ma1,main="3 plot Moving average")
ts.plot(ma2,main="5 plot Moving average")
ts.plot(ma3,main="7 plot Moving average")
ts.plot(ma3,main="9 plot Moving average")
```

We observe that as the order MA increases,the smoothing increases . However to extract the stationarity we need to use the least order MA to reduce the loss of information.

```{r }
error=data2-ma1 #
ts.plot(error)
```


Interpretation: stationarity is achieved in 3 pt moving average (MA). 
since it has a 
1)constant variance 
2)Constant mean.


Note: we need not do it for higher order to avoid loss of information. since higher order MA (lesser observation.) we tend to lose quality Information.

## Ordinary Least Square Method:

#### Concept: 


```{r }
library(readr)
LHP3 <- read_csv("C:/Users/mayur/Desktop/Mstat/Semesters/Tri-sem3/Time series/Dataset/LHP3.csv")
View(LHP3)
attach(LHP3)
```


#### Regression Model
```{r }
model=lm(LHP3$price~LHP3$month)
summary(model)
```
We observe that all the variables are statistically significant and the adjusted r^2 =0.956 shows that the model is a good fit.


```{r }
residual=resid(model)
plot(residual)
```
###Residual Analysis:

#### Normality Test

Ho: The residual follows normal distribution
H1: The residual does not follow normal distribution
```{r }
shapiro.test(residual)
```
The model does not follow normal distribution since the pvalue is less than significant alpha

## Constant Variannce

From the plot residual plot we observe that the residual in the model defies the assumption of constant variance.
This can be verified using BP test

Hypothesis:
Ho: The residual has a constant Variance
H1: The residual does not have a constantt Variance.


```{r }

library(lmtest)
bptest(model)
```
The value is less than alpha level(0.05) significance, Thus we reject null hypothesis and conclude that the residual does not follow constant variance.


### Inference: 
Since the model defies the residual analysis thus the assumptions of the error terms.
we will not get the stationary plot (stationary) to be achieved.

```{r }
ts.plot(residual)
```
The same can be verified using adf test since the data only includes trend components.

#### ADF Test

Ho: The residual is not stationary
H1: The residual is stationary

```{r }
adf.test(residual)
```
Since the P-value > o.o5 (alpha) level of significance we will fail to reject null hypothesis and conclude that the residual is not stationary.

Also, higher order of polynomial fitting is not suitable since it was observed that polynomial fitting did not significantly impact in the betterment of residual analysis.
Thus we conclude stating that Ordinary differencing method, Moving Average method were viable for the dataset to extract stationarity. However, OLS method is not suitable.


```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```


```{r }

```
